---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Loading and preprocessing the data

```{r setup}
library(plyr)
base_dir <- "C:/Users/Scott/Documents/coursera/data_science_specialization/"
course_dir <- paste0(base_dir, "reproducible_research/week_2/project_1")
setwd(course_dir)
activityRaw <- read.csv("activity.csv")
activity <- na.omit(activityRaw)
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?

```{r average_steps}
totalSteps <- ddply(activity, "date", summarize, steps = sum(steps, na.rm = TRUE))
hist(totalSteps$steps, xlab = "total steps", ylab = "counts", main = "total steps taken per day")
meanTotalSteps <- mean(totalSteps$steps)
medianTotalSteps <- median(totalSteps$steps)
```

Mean = $`r meanTotalSteps`$, median = $`r medianTotalSteps`$.

## What is the average daily activity pattern?

```{r average_pattern}
averagePattern <- ddply(activity, "interval", summarize, steps = mean(steps, na.rm = TRUE))
plot(averagePattern$steps, type = "l", xlab = "interval", ylab = "average steps", main = "average steps taken per interval")
```

Max interval (on average across all days) = `r which.max(averagePattern$steps)`.

## Imputing missing values

```{r imputing}
totalNA <- sum(is.na(activityRaw$steps))

activity <- activityRaw
activity$date <- as.Date(activity$date)
for (i in 1:nrow(activity)) {
  if (is.na(activity[i, 1])) {
    j <- which(averagePattern$interval == activity[i, 3])
    activity[i, 1] <- averagePattern[j, 2]
  }
}

totalSteps <- ddply(activity, "date", summarize, steps = sum(steps, na.rm = TRUE))
hist(totalSteps$steps, xlab = "total steps", ylab = "counts", main = "total steps taken per day")
meanTotalSteps <- mean(totalSteps$steps)
medianTotalSteps <- median(totalSteps$steps)
```

NA rows = `r totalNA`.  

Mean = $`r meanTotalSteps`$, median = $`r medianTotalSteps`$.

## Are there differences in activity patterns between weekdays and weekends?

```{r average_pattern_differences}
activity$isWeekend <- factor(weekdays(activity$date) %in% c("Saturday", "Sunday"), levels = c(FALSE, TRUE))
averagePattern <- ddply(activity, c("interval", "isWeekend"), summarize, steps = mean(steps, na.rm = TRUE))
temp = subset(averagePattern, isWeekend == TRUE)
plot(temp$steps, type = "l", xlab = "interval", ylab = "average steps")
temp = subset(averagePattern, isWeekend == FALSE)
plot(temp$steps, type = "l", xlab = "interval", ylab = "average steps")
```
